---
title: "Homework 7"
author: "Fabián A. Araneda-Baltierra"
date: "12/8/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE}
rm(list = ls())

# Load packages
library('rmutil')
library('tidyverse')
library('dplyr')
library('moments')
library('haven')
library('boot')
# Set a seed for random calculations
set.seed(11221111)
```

## Question 1

**Use data that we used for homework 4. For African American women who are currently employed, use the bootstrap to calculate the standard error of the correlation coefficient between labor wages and the workers' ages. Use 10,000 bootstrap replications. (\textit{This question also have an answer in Stata that can be found in the file HW7-FAAB.do})**

First, we load the data and then we filter our main table to get the necessary values for our analysis: gender, race, and employment status.
```{r}
# Load data
data_1 <- read.csv("ppha312x2021.csv")
data_1 <- data_1 %>% filter(incwage > 0)
data_1 <- data_1 %>% filter(sex == "Female")
data_1 <- data_1 %>% filter(race == "Black/African American/Negro")
data_1 <- data_1 %>% filter(empstat == "Employed")
```

Now we create a subset for the variables "age" and "incwage".

```{r}
subset_p1 <- data_1[,c("age", "incwage")]
subset_p1$age <- as.numeric(subset_p1$age)
```

Next we will calculate the correlation between those variables and then we will use the bootstrap to calculate the standard error of the correlation coefficient.

```{r}
corr <- function(x, indices){
  dat <- x[indices, ] 
  corr <- cor(dat[,1], dat[,2])
  return( corr )
}

coefficient_1 <- boot(subset_p1, corr, 10000)
coefficient_1
```

## Question 2
**For white men who are currently employed, use the bootstrap to calculate the standard error of the correlation coefficient between labor wages and the workers' ages. Use 10,000 bootstrap replications.**

First, we reload the data, to create a new subset, and then we filter our main table to get the necessary values for our analysis: gender, race, and employment status.
```{r}
# Load data
data_2 <- read.csv("ppha312x2021.csv")
data_2 <- data_2 %>% filter(incwage > 0)
data_2 <- data_2 %>% filter(sex == "Male")
data_2 <- data_2 %>% filter(race == "White")
data_2 <- data_2 %>% filter(empstat == "Employed")
```

Now we create a subset for the variables "age" and "incwage".

```{r}
subset_p2 <- data_2[,c("age", "incwage")]
subset_p2$age <- as.numeric(subset_p2$age)
```

Next we will calculate the correlation between those variables and then we will use the bootstrap to calculate the standard error of the correlation coefficient. We use the "corr" function previously defined.

```{r}
coefficient_2 <- boot(subset_p2, corr, 10000)
coefficient_2
```

## Question 3
**You conduct a survey that has a 10 percent refusal rate for women and a 20 percent rate for refusal rate for men. Suppose, among the respondents, the employment rate for women is 0.86 and for men it is 0.92. Derive the bounds on the employment rate:**

In general, for this problem, let's say that $x = 1$ who responded "yes" to the question "Are you employed?". Also, we are going to divide the sample, S, into people who responded $(R = 1)$ and people who did not respond $(R = 0)$. The distribution function of our sample, in this case, may be written as:

$F_S(x) = Pr(R = 1)\cdot F(x | R=1) + Pr(R = 0)\cdot F(x | R=0)$

We also know that $0 \leq F(x | R=0) \leq 1$. Then, the value of $F_S(x)$ could range over:

$Pr(R = 1)\cdot F(x | R=1) \leq F_S(x) \leq Pr(R = 1)\cdot F(x | R=1) + Pr(R = 0)\cdot 1$

We will use this last expression for the next calculations.

### A) For Woman.

$Pr(R = 1) = 0.9$

$F(x | R=1) = 0.86$

$0.9 \cdot 0.86 \leq F_S(x) \leq 0.9 \cdot 0.86 + 0.1$

$0.774 \leq F_S(x) \leq 0.874$

### B) For Man.
$Pr(R = 1) = 0.8$

$F(x | R=1) = 0.92$

$0.8 \cdot 0.92 \leq F_S(x) \leq 0.8 \cdot 0.92 + 0.2$

$0.736 \leq F_S(x) \leq 0.936$

### C) Assuming that women are half of the sample, derive the bounds to the employment rate for the combined sample of women and men.

If women are half of the sample, same happens with men. Therefore each group has a weight of $0.5$. Based on $A$ and $B$, we have the following:

$0.5 \cdot (0.774 + 0.736) \leq F_S(x) \leq 0.5 \cdot (0.874 + 0.936)$

$0.755 \leq F_S(x) \leq 0.905$


## Question 4
**The data set “Homework 7.dta” is a Stata data set of non-Hispanic respondents from the 2019 American Community Survey who reside in Virginia Beach metro area. For this exercise, use Stata only. For this exercise, use respondents who are 25 to 55 years old (inclusive). Divide people into six educational categories: less than high school including GED recipients, exactly a high school degree, some college but no degree, associates degree, exactly bachelor’s degree, more than a bachelor’s degree. Divide people into three racial groups: white, African American, and other race.**

### A) Calculate the income from wages and salaries for those with positive earnings by sex and race.
### B) Calculate the income from wages and salaries for those with positive earnings by sex and race using the weights “perwt”.

THE ANSWER TO THIS QUESTIONS CAN BE FOUND IN THE FILE: HW7-FAAB.do

## Question 5
**Use the same data set and setup as in problem 4. Create an age category variable with categories, 25-29, 30-34, 35-39, 40-44, 45-49, and 50-55. The variable ``qincwage'' is equal to four when the data are suppressed. Using our race, age, sex, and education categories along with the sample weights, construct IPW weights to account for the imputed income. Calculate income from wages and salaries by sex and race using the IPW weights. Compare these means to those in problem 4b.**

THE ANSWER TO THIS QUESTION CAN BE FOUND IN THE FILE: HW7-FAAB.do
